DATA 624 Homework 9

### 2020-04-15 DATA624/HW9.Rmd
More work on HW9

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(RColorBrewer)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

@@

```{r}
library(mlbench)
set.seed(200)
simulated <- mlbench.friedman1(200, sd = 1)
simulated <- cbind(simulated$x, simulated$y)
simulated <- as.data.frame(simulated)
colnames(simulated)[ncol(simulated)] <- "y"
```

@@

```{r}
library(randomForest)
library(caret)
rf_model_1 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_1 <- varImp(rf_model_1, scale = FALSE)
rf_imp_1
```

@@

@@

```{r}
simulated$duplicate1 <- simulated$V1 + rnorm(200) * .1
cor(simulated$duplicate1, simulated$V1)
```

@@

```{r}
rf_model_2 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_2 <- varImp(rf_model_2, scale = FALSE)
rm_imp_2
```

@@

```{r}
library(party)
rf_model_3 <- cforest(y ~ ., data = simulated)
```

@@

>>>>

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(RColorBrewer)
library(gbm)
library(randomForest)
library(caret)
library(party)
library(mlbench)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

@@

```{r}
set.seed(200)
simulated <- mlbench.friedman1(200, sd = 1)
simulated <- cbind(simulated$x, simulated$y)
simulated <- as.data.frame(simulated)
colnames(simulated)[ncol(simulated)] <- "y"
```

@@

```{r}
rf_model_1 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_1 <- varImp(rf_model_1, scale = FALSE)
rf_imp_1  %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

**No.  The scores are very close to zero indicating they are uninformative.**

@@

```{r, null_prefix = TRUE}
simulated$duplicate1 <- simulated$V1 + rnorm(200) * .1
cor(simulated$duplicate1, simulated$V1)
```

@@

```{r}
rf_model_2 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
rf_imp_2 <- varImp(rf_model_2, scale = FALSE)
rf_imp_2 %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

**The importance of `v1` has decreased significantly since adding the additional correlated predictor.  When adding the correlated predictor, both were included in the decision trees as significant predictors.  This caused the overall significance to decrease.**

@@

```{r}
rf_model_3 <- cforest(y ~ ., data = simulated)
rf_imp_3 <- varImp(rf_model_3, conditional = TRUE)
rf_imp_3 %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

@@

#### Boosted Tree

**Training it without the duplicate predictor:**

```{r}
gbm_grid <- expand.grid(.interaction.depth = seq(1, 7, by = 2),
                        .n.trees = seq(100, 1000, by = 100),
                        .shrinkage = c(0.01, 0.1),
                        .n.minobsinnode = 10)
gbm_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="gbm", tuneGrid = gbm_grid, verbose = FALSE)

gbm_imp_1 <- varImp(gbm_model_1)

gbm_imp_1$importance %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

**The boosted tree did not pick up the uninformative predictors.  Now let's see what happens when I train it WITH the duplicate predictor:**

```{r}
gbm_model_2 <- train(y ~ ., data = simulated, method="gbm", tuneGrid = gbm_grid, verbose = FALSE)

gbm_imp_2 <- varImp(gbm_model_2)

gbm_imp_2$importance %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

**This model exhibits the same pattern.  The duplicate becomes one of the important variables and the importance of `v1` decreases.**

#### Cubist

**Again training it without the duplicate predictor:**

```{r}
cubist_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="cubist")

cubist_imp_1 <- varImp(cubist_model_1)

cubist_imp_1$importance %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

**Once again, the cubist model does not pick up the un-important variables.  Let's train it WITH the duplicate predictor and see what happens:**

```{r}
cubist_model_2 <- train(y ~ ., data = simulated, method="cubist")

cubist_imp_2 <- varImp(cubist_model_2)

cubist_imp_2$importance %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(Variable = rowname) %>%
  arrange(desc(Overall)) %>%
  kable() %>%
  kable_styling()
```

**The cubist model also exhibits a similar behavior as the random forest model, but is not as pronounced.**
